---
output: pdf_document
params:
  t1: !r Sys.Date()-1
  t2: !r Sys.Date()-1
  obs: !r data.frame()
  trk: !r data.frame()
---

```{r setup, echo=FALSE}

# libraries
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rnaturalearth))

# define species color palette
obs_pal = c('right' = 'red',
             'fin' = 'blue',
             'sei' = 'lightgray', 
             'humpback' = 'green', 
             'blue' = 'darkslategrey')

# define track color palette
trk_pal = c(
  'buoy' = 'darkslategrey',
  'slocum' = 'blue',
  'plane' = '#8B6914',
  'vessel' = 'black',
  'rpas' = 'purple')

# rename inputs
t1 = params$t1
t2 = params$t2
obs = params$obs
trk = params$trk

# create data to/from message
if(t1 == t2){
  data_range = t1
} else {
  data_range = paste0(t1, ' to ', t2)
}

```

\begin{center}
	\huge \textbf{WhaleMap Summary Report} \\
	\huge \textbf{`r data_range`} \\
	\bigskip
	\large 
	\textbf{Report generated on:} `r Sys.time()`\\
	\textbf{Produced by Dalhousie MEOPAR-WHaLE project and DFO Science - Maritimes Region} \\
\end{center}

\hrulefill

#### Figure 1: Locations of large whale observations

```{r map, echo=FALSE, warning=FALSE, fig.height=8, fig.align='center'}

# combine all lat/lon to help determine plot limits
inLat = range(c(obs$lat, trk$lat),na.rm = TRUE)
inLon = range(c(obs$lon, trk$lon),na.rm = TRUE)

# get basemap data
bg = ne_countries(scale = "medium", continent = 'north america', returnclass = "sf")

# make map
map = ggplot()+
  geom_sf(data = bg)+
  coord_sf(xlim = c(-68,-51), ylim = c(41,52), expand = FALSE, )+
  # coord_sf(xlim = c(-68,-51), ylim = c(41,52), expand = FALSE)+
  geom_path(data = trk, aes(x = lon, y = lat, group = id, color = platform),
            alpha = 0.75, size = 0.3)+
  scale_color_manual(values = trk_pal)+
  geom_point(data = obs, aes(x = lon, y = lat, fill = species, size = species), 
             shape = 21, color = 'black', alpha = 0.7)+
  geom_point(data = filter(obs,species == 'right'), aes(x = lon, y = lat, fill = species, size = species), 
             shape = 21, color = 'black', alpha = 0.7)+
  scale_fill_manual(values = obs_pal)+
  scale_size_manual(values = c('right' = 3, 'fin' = 2, 'sei' = 2, 'blue' = 2, 'humpback' = 2))+
  labs(x=NULL,y=NULL,fill='Species:',size='Species:', color='Platform:')+
  theme_bw()+
  theme(panel.grid = element_blank(), 
        legend.position = 'bottom', 
        legend.box = 'vertical', 
        legend.key.height = unit(0.01,'cm'), 
        legend.spacing.y = unit(0.01,'cm'))

# show map
if(length(inLon)!=0){
  map
} 
```

```{r map_text, echo=FALSE, results='asis'}

# show map text
if(length(inLon)!=0){
  cat('For interactive map, go to: https://whalemap.org/WhaleMap/ \n')
  cat('\\clearpage \n')
} else {
  cat('*No effort or observations were recorded during this period*')
}

```

#### Table 1: Number of large whale observations

```{r spp_table, echo=FALSE, results='asis'}

# count numbers of all species sighted
vis = obs[obs$score=='definite visual',]
vis = obs[!is.na(obs$number),]

if(nrow(vis)!=0){
  vis = aggregate(number ~ species, data = vis, FUN = sum)
} else {
  vis = data.frame(species = levels(obs$species), number = rep(0, length(levels(obs$species))))
}

# count acoustic detections
aco = as.data.frame(table(obs$species[obs$score=='definite acoustic']))

# merge sightings and acoustics
spp_table = merge(vis, aco, by.x = 'species', by.y = 'Var1', all.y = TRUE)

# rename columns
colnames(spp_table) = c('species', 'visual_counts', 'acoustic_detections')

# replace zeros
spp_table$visual_counts[is.na(spp_table$visual_counts)] = 0

# sort table alphabetically
spp_table=spp_table[order(as.character(spp_table$species), decreasing = FALSE),]

# display table
kable(spp_table, row.names = FALSE, align = 'c')  

```

*Visual counts only include sightings where numbers of whales were recorded, and may include duplicates*

#### Table 2. Platforms with recorded survey effort

```{r effort, echo=FALSE, results='asis'}

# display table or message
if(nrow(trk)==0){
  cat('*No platforms recorded effort during this period*')
} else {
  platform_table = aggregate(platform ~ name, data = trk, FUN = unique)
  kable(platform_table, row.names = FALSE, align = 'c')
}

```

#### Table 3: North Atlantic right whale sightings

```{r narw_sig, echo=FALSE, results='asis'}

# select species
sig = obs[obs$species == 'right' & obs$score == 'definite visual',]

# format table
sig = sig[,c('date','time', 'lat', 'lon', 'number', 'calves', 'platform', 'name')]

# sort by time
sig = sig[order(sig$time),]

# format timestamps
sig$time = format(sig$time, '%H:%M:%S')

# display table or message
if(nrow(sig)==0){
  cat('*No right whales were sighted*')
} else {
  kable(sig, row.names = FALSE, align = 'c')  
}

```

#### Table 4: North Atlantic right whale acoustic detections

```{r narw_det, echo=FALSE, results='asis'}

# select species
det = obs[obs$species == 'right' & obs$score == 'definite acoustic',]

# format table
det = det[,c('date','time', 'lat', 'lon', 'platform', 'name')]

# sort by time
det = det[order(det$time),]

# format timestamps
det$time = format(det$time, '%H:%M:%S')

# display table or message
if(nrow(det)==0){
  cat('*No right whales were acoustically detected*')
} else {
  kable(det, row.names = FALSE, align = 'c')  
}

```
